Dense depth estimation from image pairs

### 2021-06-30 paper/paper.tex
nice

Let $C_1$ and $C_2$ denote the two cameras, $I_1$ and $I_2$ denote the two images, and let $h$ be the estimated depth map for $C_1$.
Our fully non-linear cost function is
\begin{equation}
\begin{split}
% E(h) &= \lambda \int_0^1\int_0^1{\Omega_0}\left\|I_1(\pi(\exp(\hat{\eta}_1)X(x,h)) - I_0(\pi(x))\right\|
%      + \|\nabla h\| \,dx\,dy,
E(h) &= \lambda \frac{1}{|\Omega^*|}\int_{\Omega^*}\left\|I_2(\text{reproj}(x,y)) - I_1(x,y)\right\|\,dx\,dy \\
     &+ \int_0^1\int_0^1\|\nabla h\| \,dx\,dy.
\end{split}
\end{equation}
which has TV-$L^1$ \cite{tv_optical_flow} data and regularizer terms. $\text{reproj}(x,y)$ is where the non-linear complexity is.
$x$ and $y$ denote image coordinates in $I_1$, and $h$ gives a transformation from these image coordinates to world-space,
given the camera pose of $C_1$. We can then project this point to the image space of $I_2$. $\Omega^* \subset [0,1]^2$
is defined as
\begin{equation}
    \Omega_0 \coloneqq \left\{(x,y) \in [0,1]^2 \mid \text{reproj}(x,y) \in [0,1]^2\right\}.
\end{equation}
Therefore, our data term is weighted over those pixels which have valid reprojections. Without the $1/|\Omega^*|$ factor
in the data term, our optimization would likely cause $h$ to be a constant function (giving zero regularizer cost), large enough such that
all reprojections are invalid (giving zero data cost) --- Weighting by $1/|\Omega^*|$ has been effective in our implementation.

Let $\rho(h) = \left\|I_1(\pi(\exp(\hat{\eta}_1)X(x,h)) - I_0(\pi(x))\right\|$. Since the data term is non-quadratic, we
cannot apply the ROF optimization method directly. However, we can introduce
an auxiliary depth map $h^\prime$ to separate the data and regularizer terms,
and introduce a penalizer for disparity between $h^\prime$ and $h$. The
new separable cost function is
\begin{equation}\label{separable_cost}
    \hat{E}_\theta(h, \hat{h}) = \int_\Omega \lambda \|\rho(h)\| + \frac{1}{2\theta}(h - \hat{h})^2 + \|\nabla \hat{h}\|.
\end{equation}
The parameter $\theta$ is set to some small constant.
The cost function $E_\theta$ is separated into functions of $h$ and $h^\prime$,
and $h$ and $h^\prime$ are alternately optimized in the following way:
\begin{itemize}
    \item For $h$ fixed, solve
        \begin{equation}\label{minh}
            \min_{h^\prime} \int_\Omega \lambda \|\nabla h^\prime\| + \frac{1}{2\theta}(h - h^\prime)^2\,dx
        \end{equation}
    using the ROF optimization methods.
    \item For $h^\prime$ fixed, solve
        \begin{equation}\label{minhhat}
            \min_h \int_\Omega \|\rho(h)\| + \frac{1}{2\theta}(h - h^\prime)^2\,dx.
        \end{equation}
        Since $\rho(h)$ measures point-wise reprojection error, this
        optimization can be done point-wise.
\end{itemize}
The ROF cost function \eqref{minh} can be minimized using the iterative method of Chambolle \cite{chambolle}. For completeness, we reproduce this
method here in the context of dense depth map estimation. We provide no proofs --- see \cite{chambolle} for full details.

>>>>

\subsection{The optimization problem}
Let $C_1$ and $C_2$ denote the two cameras, $I_1$ and $I_2$ denote the two images, and let $h$ be the estimated depth map for $C_1$.
Our fully non-linear cost function is
\begin{equation}\label{nonlinear_cost}
\begin{split}
% E(h) &= \lambda \int_0^1\int_0^1{\Omega_0}\left\|I_1(\pi(\exp(\hat{\eta}_1)X(x,h)) - I_0(\pi(x))\right\|
%      + \|\nabla h\| \,dx\,dy,
E(h) &= \frac{\lambda}{|\Omega^*|}\int_{\Omega^*}\left\|I_2(\text{reproj}(x,y)) - I_1(x,y)\right\|\,dx\,dy \\
     &+ \int_0^1\int_0^1\|\nabla h\| \,dx\,dy.
\end{split}
\end{equation}
which has TV-$L^1$ \cite{tv_optical_flow} data and regularizer terms. $\text{reproj}(x,y)$ is where the non-linear complexity is.
$x$ and $y$ denote image coordinates in $I_1$, and $h$ is used to transform these image coordinates to world-space,
given the camera pose of $C_1$. We can then project this point to the image space of $I_2$. $\Omega^* \subset [0,1]^2$
is defined as
\begin{equation}
    \Omega_0 \coloneqq \left\{(x,y) \in [0,1]^2 \mid \text{reproj}(x,y) \in [0,1]^2\right\}.
\end{equation}
Therefore, our data term is weighted over those pixels which have valid reprojections. Without the $1/|\Omega^*|$ factor
in the data term, our optimization would likely cause $h$ to be a constant function (giving zero regularizer cost), large enough such that
all reprojections are invalid (giving zero data cost) --- In our implementation, weighting by $1/|\Omega^*|$ has been effective at preventing this.

\subsection{Splitting the cost function}
Let 
$$\rho(h) = \left\|I_2(\text{reproj}(x,y)) - I_1(x,y)\right\|.$$
Since the data term is non-quadratic, we
cannot apply the ROF optimization method directly. However, we can introduce
an auxiliary depth map $h^\prime$ to separate the data and regularizer terms,
and introduce a penalizer for disparity between $h^\prime$ and $h$ (the same method as used in \cite{tv_optical_flow} for optical flow). The
new separable cost function is
\begin{equation}\label{separable_cost}
    \hat{E}_\theta(h, h^\prime) = \int_\Omega \lambda \|\rho(h)\| + \frac{1}{2\theta}(h - h^\prime)^2 + \|\nabla h^\prime\|.
\end{equation}
The parameter $\theta$ is set to some small constant.
The cost function $E_\theta$ is separated into functions of $h$ and $h^\prime$,
and $h$ and $h^\prime$ are alternately optimized in the following way:
\begin{itemize}
    \item For $h$ fixed, solve
        \begin{equation}\label{minh}
            \min_{h^\prime} \int_\Omega \lambda \|\nabla h^\prime\| + \frac{1}{2\theta}(h - h^\prime)^2\,dx
        \end{equation}
    using the ROF optimization methods.
    \item For $h^\prime$ fixed, solve
        \begin{equation}\label{minhhat}
            \min_h \int_\Omega \|\rho(h)\| + \frac{1}{2\theta}(h - h^\prime)^2\,dx.
        \end{equation}
        Since $\rho(h)$ measures point-wise reprojection error, this
        optimization can be done point-wise.
\end{itemize}

\subsection{Efficiently solving the ROF step}
The ROF cost function \eqref{minh} can be minimized using the iterative method of Chambolle \cite{chambolle}. For completeness, we reproduce this
method here in the context of dense depth map estimation. We provide no proofs --- see \cite{chambolle} for full details.

\subsection{Interpretation of the algorithm and its parameters}
We are left with many parameters: regularizer parameter $\lambda$, thresholding parameter $\theta$, chosen each iteration, and the number of iterations.
The performance of our method relies heavily on the choices of these values --- the preliminary results seen in figure \ref{cloud_computed}
were only achieved by manual adjustment. Although, in principle, we aim to globally optimize the original cost function \eqref{nonlinear_cost},
our final method of optimization can be reinterpreted as a step-based algorithm:
\begin{itemize}
    \item
\end{itemize}
